---
title: "OMIS 482 Project Part 1"
author:"Gary Dhami, Francisco Lozano, Amelasky Mendez, Tyler Williams, Rachel Worden"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: word_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction
Hi, we are a group of movie reviewers! We pick a movie every month, watch it, and get together to discuss it. We love the movies so much that we wanted to ask what makes a certain movie successful? 

Using the movies dataset below, our group plans to analyze what factors affect the success of a movie. We believe a movie is successful by how much revenue a movie makes. By analyzing revenue, we can determine success and get a better idea of what other variables make a movie a hit. Determining this would allow movie producers and companies to know what they can do to maximize success for their next film. The analysis is done using data from an excel file and will be manipulated and visualized using tidyverse in Rstudio. 

## Importing the data:
```{r, echo=F}
library(tidyverse)
Movies_Data <- read_csv("movies_project.csv")
print(Movies_Data)
```

## Tidying the data:
Before we can manipulate the dataset and start to analyze the data, we need to tidy it up! We noticed that the columns "keywords", "index", "overview", "homepage", "ID", "Tagline","runtime" and "Cast" were not relevant, so we decided to omit these columns because they add no value to our goal of trying to figure out what impacts a movie's success. As of right now our knowledge in Rstudio doesn't allow us to perform analytics on text columns and these categories are mostly character/text columns. In order to find the answers to our questions we plan to use mostly integer values. We want to make sure we abide by "rule of 3" ensuring each variable has its own column, row, cell. In separating release_date we left the resulting columns to data type chr because month, day, and year aren't columns that can be summarized with mean,max,sum, and so forth. 

## Filtering:
Although they are columns that can be used to group by. Next, since our data set is not in the tidyverse package, we had to assign our data frame to the name "movies." Because our dataset goes all the way back to 1930 and even predicts films to 2029, we found that using our data set from 1980 to 2017 would give us a more accurate look at how film has evolved. We also found that in order to measure a movies success it has to be released, so we filtered out the movies that were never released. Finally, we removed all the outliers in the data set. Most of the outliers were because the movies were really popular and broke records. For example, Avatar was a popular movie that revolutionized the movie industry since it used a bunch of new technology such as new motion capture animation technologies. 


##Summary Statistics - Rachel Worden
Descriptive statistics are a vital part in interpreting a dataset. It allows the reader and user(s) of the code to get an understanding of the data. In this case, we are taking a very large number of observations and turning it into something simple. The descriptive statstics below can help movie production companies estimate if they are below, the same, or above industry standard when it comes to revenue. For example the entertainment conglomorate, Disney could look at the average in order to estimate how much money their next film should be able to bring in. Once we have the code in order to compute these statistics, we can then group by other columns we want to analyze. THis will allow us to see if there are correlations between revenue and other variables. Out of 3,376 observations in our data set, the average revenue a movie makes is 117,031,353. With a wide standard deviaiton of 183483090. 

movies %>% 
  filter(!is.na(revenue)) %>% 
  filter(revenue > 0) %>%
  summarise(count=n(),avg_rev=mean(revenue),median_rv=median(revenue),max_rev=max(revenue),min_rev=min(revenue),sd_rev=sd(revenue))

movies %>%  
  group_by(genres) %>% 
  filter(!is.na(revenue)) %>% 
  summarise(count=n(),avg_rev=mean(revenue),median_rv=median(revenue),max_rev=max(revenue),min_rev=min(revenue),sd_rev=sd(revenue, na.rm = T))




movies <-select(Movies_Data,2,3,7:8,10:14,16:18,20:22) %>%
  filter(!is.na(revenue),!is.na(release_date)) %>%
 separate(director,into = c("director_first_name","director_last_name"), sep = " ",convert = "TRUE") %>%
 separate(release_date, into = c("month", "day", "year"), sep = "/") %>%  filter(year>= 1980 & year <= 2017,status=="Released",original_title != "Avatar",original_title != "Titanic")


```
## Analysis 1: Does Language effect Revenue? - Francisco Lozano
We wanted to see if language effects revenue, so we grouped the data set by the column original_language. After the data set was grouped we ran descriptive statistics such as mean, median, max, min, and standard deviation. Finally we plotted the average of the revenue on a scatter plot. From the graph we saw that English and Tegulu had the highest averaged revenue. We first hypothesized that English would be the first, so we were surprised Tegulu had a high average revenue. Although we then noticed that the size of the point was small meaning that not a lot movies were made in that language. We then made another graph that plotted the movies revenue based on original language to dive deeper into the data. From there we noticed the data was skewed because Tegulo only had one movie made in that language resulting in the average to be high. While English had the most movies made in that language. We also saw that English did have a higher influence in revenue since all of the top movies based on revenue were originality made in English. Confirming that language does effect revenue and the success of a movie.
```{r}
movies %>%
group_by(original_language) %>%
summarise(count=n(),rev_mean = mean(revenue,na.rm = TRUE),rev_med=median(revenue),rev_max=max(revenue),
          rev_min=min(revenue),rev_sd=sd(revenue)) %>%
ggplot() +
geom_point(mapping = aes(y = original_language, x=rev_mean,size=count))

movies %>%
ggplot() +
geom_point(mapping = aes(y = original_language, x=revenue))
```

## Analysis 2: Does Country of production have an effect on revenue?- Amelasky Mendez
We wanted to see if the country production of a movie influences the revenue it generates. We started by grouping the data by country_production. We then ran descriptive statistics on the data that would measure the Min, Max , mean, median, and standard deviation. With these results we created a scatter plot that helped us visualize the mean amount of revenue the movies generated in each country of production. We believed that the US was going to be one of the country of productions to have high average revenue. However, it is not even in the top 5, DM had the highest average revenue followed by MT, CZ, AE, and MA. Judging by the scatter plot point size they differ from that of the US. By running another plot, we can see that the US has produced more movies than all the top 5 countries combined which created an unbalanced data in the previous plot. This shows that the place of production does have an influence on the amount of revenue a movie generates.


```{r}
movies%>%
  group_by(country_production)%>%
  summarise(count=n(), rev_min=min(revenue), rev_max=max(revenue),rev_mean=mean(revenue), rev_med=median(revenue),rev_sd=sd(revenue))%>%
  
  ggplot()+
  geom_point(mapping=aes(y=country_production, x=rev_mean, size=count))
  movies%>%
  ggplot()+
  geom_point(mapping = aes(y=country_production, x=revenue))
```

## Analysis 3: Does the movie popularity have an effect on Revenue? - Gary Dhami
We wanted to see if the popularity of a movie effects the revenue, so we grouped the data set by the column popularity. After the data set was grouped we ran descriptive statistics such as mean, median, max, and min. We can see in the tibble that the movie with the highest revenue had popularity of 144. Finally we plotted the revenue on a scatter plot. Our assumption was that the higher the popularity, the higher the revenue for the movie which does not hold true. From the graph we can see there is no real correlation between movies that have higher popularity or have higher revenues. The graph displays that the movies with the higher revenues can still have popularity less than 250 and movies with poularity higher than 250 can still have low revenues. The movie popularity has minimal effect on movie revenue. 

```{r}
movies %>%
group_by(popularity) %>%
summarise(count=n(),rev_mean = mean(revenue,na.rm = TRUE),rev_med=median(revenue),rev_max=max(revenue),
rev_min=min(revenue)) %>% arrange(desc(rev_mean))

movies %>%
group_by(popularity) %>%
summarise(count=n(),rev_mean = mean(revenue,na.rm = TRUE),rev_med=median(revenue),rev_max=max(revenue), 
rev_min=min(revenue)) %>% 
arrange(desc(rev_mean)) %>% ggplot() + 
geom_point(mapping = aes(x = popularity, y=rev_mean))
```


## Analysis 4: Does the movie rating have an effect on Revenue? - Gary Dhami
We wanted to see if the movie rating of a movie effects the revenue. So we grouped the data set by the column vote_average. After the data set was grouped we ran descriptive statistics such as mean, median, max, and min. We can see in the tibble that the movie with the higher vote_avergae did have the higher revenue.We plotted the revenue on a scatter plot and the graph displays a relationship between movie ratign and revenue. Our assumption was that the higher the movie rating, the higher the revenue for the movie which does hold true. we also plotted a line graph that does a great job displaying how the revenue goes up for data points with higher movie ratings. The two graphs agree with each other, leading us to believe that movie ratings has an effect on movie revenue. 

```{r}
movies %>%
group_by(vote_average) %>% filter(revenue > 0) %>% 
summarise(count=n(),rev_mean = mean(revenue,na.rm = TRUE),rev_med=median(revenue),rev_max=max(revenue), rev_min=min(revenue)) %>% arrange(desc(rev_mean))

movies %>%
group_by(vote_average) %>% filter(revenue > 0) %>% 
summarise(count=n(),rev_mean = mean(revenue,na.rm = TRUE),rev_med=median(revenue),rev_max=max(revenue), rev_min=min(revenue)) %>% arrange(desc(rev_mean)) %>% ggplot(data=movies, mapping=aes(x=vote_average, y=revenue))+geom_point()+geom_smooth()

movies %>%group_by(vote_average) %>% filter(revenue > 0) %>% 
ggplot() + 
geom_smooth(mapping = aes(y = revenue, x=vote_average))
```
## Analysis 5: Does the month a movie is released effect the Revenue? Tyler Williams
To see if the month of release effects the revenue, we first looked at some descriptive statistics for revenue grouped by month. By looking at the top average revenues by month, we see that June has the highest average revenue. We then created a scatter plot with revenue on the y-axis and release month on the x-axis to see the trend between month and revenue. The results show that some months are better to release a movie than others. Late sping to early summer produces movies with higher revenues often. The holiday months of Novemeber and Decemeber also see higher revenues. The winter and early fall generally are bad times to release a movie, as revenues are frequently lower than movies released in more optimal months
```{r}
movies %>% 
  group_by(month) %>% 
  summarise(count=n(),rev_mean = mean(revenue,na.rm = TRUE),rev_med=median(revenue),rev_max=max(revenue),
          rev_min=min(revenue),rev_sd=sd(revenue))  %>% 
  arrange(desc(rev_mean))
movies %>%
  ggplot() +
  geom_point(mapping = aes(x=month,y=revenue))
```
## Analysis 6: Does the budget have an effect on revenue? - Rachel Worden
When analyzing a movies successs, revenue and budget seem to be synonomous. However, just because a movie production company might have a large bugdget, doesn't necessairly mean they will make tons of revenue. Some movies have done very well without having to invest huge sums of money. To test if there is a correlation between revenue and budget we can create descriptive statstics, by grouping by budget and measuring revenue. {FILL IN} Then.....plot them on a scatter plot to see if there is a relationship between the two. 


movies %>%  
  group_by(budget) %>% 
  filter(!is.na(revenue)) %>% 
  filter(revenue > 0) %>%
  filter(budget >0)  %>%
  summarise(count=n(),avg_rev=mean(revenue),median_rv=median(revenue),max_rev=max(revenue),min_rev=min(revenue),sd_rev=sd(revenue)) 
  
  movies %>% 
  filter(!is.na(budget)) %>% 
  summarise(count=n(),avg_budget=mean(budget),median_budget=median(budget),max_budget=max(budget),min_buget=min(budget),sd_rev=sd(budget))
  

ggplot(movies)+
  geom_point(mapping = aes(x=revenue, y=budget, linetype=_____)




```{r}


## Ideas:
In the column "status" filter by released only 
Status column look for correlation between released and months -stick with one variable ( revenue)
Min, max, mean, standard deviation, of revenue 
language and country 
popularity(?)

how does it fit in the overall idea(research question)

```
```
## Questions we have:

Is there a correlation between budget and the amount of revenue a movie makes? 
Is there a correlation between 
What is the most popular genre? 
Is there a correlation between popularity and release date
Is there a correlation between how successful a movie is and their production company? 
Is there a correlation between movie ratings and revenue?


```

```

```



```

```
